### 2020-03-03 poster/src/algorithms.tex
Poster

\documentclass[11pt,twoside,a4paper]{article}

\usepackage[ruled, vlined, linesnumbered]{algorithm2e}
\usepackage{algcompatible}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{changepage}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{fancyhdr} % Header
\usepackage[margin=1.4in]{geometry}

\pagenumbering{arabic}

\algblockdefx{FORALLP}{ENDFAP}[1]%
  {\textbf{for all }#1 \textbf{do in parallel}}%
  {\textbf{end for}}

\begin{document}

\begin{algorithm}
\caption{Na\"ive Single-Pass Insertion-Streaming Algorithm for Neighbourhood Detection}
\SetKwInOut{Require}{require}
\Require{Stream $\{(s_0,t_0)\dots(s_n,t_n)\}$, degree bound $d$, precision bound $c$}
$N\leftarrow\{\{\}\}\ \{\text{neighbourhoods}\}$\\
\For{$i=0\dots n$} {
	append $t_i$ to $N[s_i]$\\
	\lIf{\text{size}($N[s_i])= \frac{d}{c}$} {
		\Return{($s_i$,$N[s_i]$)}
	}
}
\Return{FAIL}
\end{algorithm}

\begin{algorithm}
\caption{One-pass $c$-approximation Insertion-Only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
\SetKwInOut{Require}{require}
\Require{Space $s$, degree bound $d$.}
$s\leftarrow\lceil\ln(n)\cdot n^{\frac1c}\rceil$\\
\For{$i\in[0,c-1]$\textbf{ in parallel}}{
	$(a_i,S_i)\leftarrow\text{Deg-Res-Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
}
\Return{Uniform random neighbourhood $(a_i,S_i)$ from the successful runs}
\end{algorithm}

\begin{algorithm}
\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Vertex Sampling)}}
\SetKwInOut{Require}{require}
\Require{Space $s$, degree bound $d$.}
Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$\\
Sample a uniform random subset $A'\subseteq A$ of size $10\ x\ln n$ of vertices.\\
\For{$a\in A'$}{
Run $10\frac{d}{c}\ln n$ $l_0$-samplers on the set of edges incident to $a$.
}
\Return{Any neghbourhood of size $\frac{d}{c}$ among the stored edges, if there is one.}
\end{algorithm}

\begin{algorithm}
\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Edge Sampling)}}
\SetKwInOut{Require}{require}
\Require{Space $s$, degree bound $d$.}
Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$\\
Run $20\frac{nd}{c}\left(\frac1x+\frac1c\right)\ln(n)$ $l_0$ samplers on the stream, in parallel. Each producing a uniform random edge.\\
\Return{Any neghbourhood of size $\frac{d}{c}$ among the stored edges, if there is one.}
\end{algorithm}

\end{document}